Extending the WebID Protocol with Access Delegation

### 2012-07-15 paper.tex

\author{Sebastian Tramp\inst{1} \and Henry Story\inst{2} \and Philipp Frischmuth\inst{1} \and Timofey Ermilov\inst{1} \and S\"oren Auer\inst{1}}

\institute{
Universit\"at Leipzig, Institut f\"ur Informatik, AKSW,\\
Postfach 100920, D-04009 Leipzig, Germany,\\
\email{\{lastname\}@informatik.uni-leipzig.de}\\
\url{http://aksw.org/FirstnameLastname} (WebID)
\medskip\and
Apache Foundation\\ 
\email{henry.story@bblfish.net}\\
\url{http://bblfish.net/people/henry/card\#me} (WebID)
}

@@

In the context of access delegation, we distinguish a WebID which refers to an agent which has direct access to a certain resource from a WebID which refers to an agent which has access to a resource only if he acts in the name of the first agent.
We call the latter agent the \textit{secretary agent} and the first one the \textit{named agent}.
\todo{maybe another name here?}

@@

\subsection{Basic Use-Cases for Access Delegation}\label{sec:usecases}

@@

Currently, the default usage intention for WebIDs are browser-based authentication scenarios.
This is an important use-case and should be of special interest but we use the world wide web not only from our browsers but also from other applications and from other application on other devices such as smartphones, tablets and freedomboxes\footnote{\url{https://freedomboxfoundation.org/}}.
In a typical every day usage scenario, we use feed readers and social network clients hand in hand with web browsers.
This is true esp. for mobile devices where specialized clients for social activities are widely in use and does exist even for academically driven social networks as the distributed Social Web based on FOAF WebIDs~\cite{tramp-s-2011--a}.

@@

\paragraph{Working Groups}\label{sec:WorkingGroups}

Access for working groups to a resource is typically specified in a way that each resource guard needs to hold a list of group members of this group.
In the Social Semantic Web, these working groups have its own WebID resources which describe the group as an entity and its members as relations to their user WebIDs.
To authorize a group of agents to access a resource should be as simple as to a single agent.
In order to allow this, we define the group as the named agent and all group members as its secretary agents.

@@

In this context, the group is the named agent and all current group members are secretary agents.

\paragraph{Office Roles}

An office role is a construct where a user can access specific resources because of his working description.
A product manager has access to all product related resources because of her role as product manager and not because of her identity.
The role itself is identified by WebID and all access authorization rules should use this role WebID instead of a user WebID.

@@

\lstinputlisting[
    name=schema,style=turtle,float=htb,label=listing:schema,
    basicstyle=\ttfamily\scriptsize,
    caption={Schema description of the delegation property},
]{Listings/schema.ttl}

>>>>

\author{Sebastian Tramp\inst{1} \and Henry Story\inst{2} \and Philipp Frischmuth\inst{1} \and Timofey Ermilov\inst{1} \and S\"oren Auer\inst{1} \and Andrei Sambra\inst{3}}

\institute{
Universit\"at Leipzig, Institut f\"ur Informatik, AKSW,\\
Postfach 100920, D-04009 Leipzig, Germany,\\
\email{\{lastname\}@informatik.uni-leipzig.de}\\
\url{http://aksw.org/FirstnameLastname} (WebID)
\medskip\and
Apache Foundation\\ 
\email{henry.story@bblfish.net}\\
\url{http://bblfish.net/people/henry/card\#me} (WebID)
\medskip\and
CNRS Samovar UMR 5157, TELECOM SudParis\\
\email{andrei.sambra@it-sudparis.eu}\\
\url{https://my-profile.eu/people/deiu/card\#me} (WebID)
}

@@

In the context of access delegation, we distinguish a WebID which refers to an agent which has direct access to a certain resource from a WebID which refers to an agent which has access to a resource only if he acts in the name of the first agent.
We call the latter agent the \textit{secretary agent} and the first one the \textit{named agent}.
\todo{maybe another name here? (A:Agent and delegatedAgent maybe?)}

@@

\subsection{Basic Use-Cases for Access Delegation}\label{sec:usecases}\todo{A: IMHO two major cases, 'attended' and 'unattended'}

@@

Currently, the default usage intention for WebIDs are browser-based authentication scenarios.
This is an important use-case and should be of special interest but we use the world wide web not only from our browsers but also from other applications and from other application on other devices such as smartphones, tablets and freedomboxes\footnote{\url{https://freedomboxfoundation.org/}}.
In a typical every day usage scenario, we use feed readers and social network clients hand in hand with web browsers.
This is true esp. for mobile devices where specialized clients for social activities are widely in use and does exist even for academically driven social networks as the distributed Social Web \todo{A: req. explaining? (Web2.0 vs 3.0 ambiguity)} %Web2.0 or 3.0? the begining of the sentence talks about 2.0, and you end with FOAF (3.0)?
 based on FOAF WebIDs~\cite{tramp-s-2011--a}.

@@

\paragraph{Working Groups}\label{sec:WorkingGroups}
\todo{A: why 'working' and not just 'Groups'? (difference?)}
Access for working groups to a resource is typically specified in a way that each resource guard \todo{A: is it really the guard who holds the list, or some access control sys?} needs to hold a list of group members of this group.
In the Social Semantic Web \todo{A: see my prev. note about Web2.0 vs 3.0}, these working groups have its own WebID resources which describe the group as an entity and its members as relations to their user WebIDs.
To authorize a group of agents to access a resource should be as simple as to a single agent.
In order to allow this, we define the group as the named agent and all group members as its secretary agents.

@@

In this context, the group is the named agent and all current group members are secretary agents.\todo{A: not sure I agree (see source)}%the group is not an entity, therefore it cannot take actions by iteself. Actions are always performed by users, as part of the group.

\paragraph{Office Roles}
\todo{A: how is this different from the prev. case?}
An office role is a construct where a user can access specific resources because of his working description.
A product manager has access to all product related resources because of her role as product manager and not because of her identity.
The role itself is identified by WebID and all access authorization rules should use this role WebID instead of a user WebID.

@@

\lstinputlisting[
    name=schema,style=turtle,float=htb,label=listing:schema,
    basicstyle=\ttfamily\scriptsize,
    caption={Schema description of the delegation property},
]{Listings/schema.ttl}\todo{A:why add the public key of the agent here? (revocation issues)}
